---
title: "State space modeller"
author: "Mia Risbjerg Steiner"
date: "8/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This exercise is about the state space models and the Kalman filter as discussed
in Mathematical Modelling. During the oral
exam you will have 20 min to present the exercise. You decide what topics to
cover and how to present them, however, we will ask questions to any part of the
exercise and presentation. Feel free to work in small groups.

## Kalman filter

Consider the dynamic linear model in the general form given in Section
2.4 of the book ``Dynamic Linear Models with R''. Go through the
Kalman filter (and, if time permits, the Kalman smoother) for the
simple random walk + noise model. Interpret the steps.

```{r}
hej <- c(1,2,3)
```

## Modelling daily milk yield

Focus on the milk yield data, lecture and R-script of session 2. Feel
free to elaborate on the script.  In the script we focus on milk yield
of one cow (and we extract prior information from a bunch of other
cows).

We have several competing modelling tools at hand

1. Random walk + noise model

1. Local linear growth model

1. The parametric woods curve and a dynamic regression model

1. The parametric woods curve and a Bayesian linear model

1. The parametric woods curve and a classical linear model (as you
   studied on 5th semester).

Things to look at (take a look at the initial R-script): In the
R-script focus is on prediction for one cow, and you may want to keep
focusing on this cow.

1. Given data up to day 20, up to day 50 and up to day 100 forecast
  the yield in the rest of the lactation period for each of the five
  models for this specific cow.

1. How will you asses if a model is good model? Which model seems
   best?

1. Do the same on another cow; are your conclusions the same?

1. For some (all?) models you need prior distributions on some
   parameters.  Feel free to follow the approach taken in the
   R-script, although you may want to rethink the choice of batch of
   cows from which the prior is parameters are "extracted".

1. For some (all?) models there are unknown parameters to be
   estimated. Estimate these parameters. The question is on the basis
   of which data these parameters should be estimated? The batch of
   cows from which priors were extracted? The cows you want to make
   forecast about? Another batch of data? Make a choice and argue your
   case.